---
title: "R Notebook"
output: html_notebook
---

Load libraries

```{r}
#BiocManager::install("AnVILGCP")
#BiocManager::install("ggVennDiagram")
library(AnVIL)
library(tidyverse)
library(ggVennDiagram)
library(RColorBrewer)
```

Read data from tables

```{r}
participant <- avtable("participant")
phenotype <- avtable("phenotype")
experiment <- avtable("experiment")
experiment_dna_short_read <- avtable("experiment_dna_short_read")
genetic_findings <- avtable("genetic_findings")
nrow(participant)
```

```{r}
for (t in unique(experiment$table_name)) {
  experiment[[paste0(t, "_id")]] <- ifelse(experiment[["table_name"]] == t, experiment[["id_in_table"]], NA)
}

experiments <- experiment %>%
  left_join(select(experiment_dna_short_read, experiment_dna_short_read_id, experiment_type)) %>%
  mutate(data_type = sub("^experiment_", "", table_name)) %>%
  mutate(data_type = ifelse(!is.na(experiment_type), paste(data_type, experiment_type, sep="_"), data_type)) %>%
  select(participant_id, data_type)

# remove families with multiple probands
duplicates <- participant %>%
  filter(proband_relationship == "Self") %>%
  count(family_id) %>%
  filter(n > 1)

participant <- participant %>%
  filter(!(family_id %in% duplicates$family_id))

participants <- participant %>%
  select(participant_id, family_id, proband_relationship) %>%
  inner_join(experiments)

families <- participants %>%
  group_by(family_id) %>%
  count(data_type)

probands <- participants %>%
  filter(proband_relationship == "Self") %>%
  group_by(family_id) %>%
  count(data_type)
```


# How many families have multi-omics data?

```{r}
data_types <- list(
  "ATAC" = c("atac_short_read"),
  "DNA short read" = c("dna_short_read_exome", "dna_short_read_genome"), 
  "DNA long read" = c("nanopore", "pac_bio"),
  "RNA" = c("rna_short_read")
)
x <- lapply(data_types, function(t) {
  probands %>%
    filter(data_type %in% t) %>%
    select(family_id) %>%
    unlist(use.names=FALSE)
})

ggVennDiagram(x) + scale_fill_gradient(low="grey90", high="orange")
```


```{r}
data_types <- list(
  "SR exome" = c("dna_short_read_exome"), 
  "SR genome" = c("dna_short_read_genome"), 
  "Nanopore" = c("nanopore"), 
  "PacBio"= c("pac_bio")
)
x <- lapply(data_types, function(t) {
  probands %>%
    filter(data_type %in% t) %>%
    select(family_id) %>%
    unlist(use.names=FALSE)
})

ggVennDiagram(x) + scale_fill_gradient(low="grey90", high="orange")
```



How many families have at least one individual with multi-omic data (proband or not)?

```{r}
tmp <- participants %>%
  group_by(family_id, participant_id) %>%
  summarise(n_data_types=n()) %>%
  filter(n_data_types > 1) %>%
  group_by(family_id) %>%
  summarise(n_multi_omic=n())
nrow(tmp)
```

How many families have multiple data types for > 1 family member?

```{r}
tmp2 <- tmp %>%
  filter(n_multi_omic > 1)
nrow(tmp2)
```

For all the probands with multi-omic data, how many have family members also with multi-omic data?

```{r}
mo <- participants %>%
  group_by(family_id, participant_id, proband_relationship) %>%
  summarise(n_data_types=n()) %>%
  filter(n_data_types > 1)

pb_mo <- mo %>%
  filter(proband_relationship == "Self")

fam_mo <- mo %>%
  filter(proband_relationship != "Self") %>%
  filter(family_id %in% pb_mo$family_id) %>%
  ungroup() %>%
  distinct(family_id)
nrow(fam_mo)
```



# Solve status

```{r}
solve <- participant %>%
  filter(proband_relationship == "Self", solve_status != "Unaffected") %>%
  mutate(solve_status = factor(solve_status, levels=c("Solved", "Probably solved", "Partially solved", "Unsolved"))) %>%
  select(participant_id, family_id, solve_status)

count(solve, solve_status)
```


# Genetic findings

```{r}
sorted <- function(x,y,z) {
  st <- paste(sort(na.omit(c(x,y,z))), collapse=" | ")
  if (st == "") st <- NA
  return(st)
}
findings <- solve %>%
  left_join(genetic_findings) %>%
  mutate(variant_type = ifelse(variant_type %in% c("SNV", "INDEL"), "SNV/INDEL", variant_type)) %>%
  separate_wider_delim(method_of_discovery, delim="|", names=paste0("method", 1:3), too_few="align_start") %>%
  mutate(across(starts_with("method"), str_trim)) %>%
  rowwise() %>%
  mutate(method_of_discovery = sorted(method1, method2, method3))

count(findings, solve_status, method_of_discovery, variant_type)
```

```{r}
methods <- unique(na.omit(findings$method_of_discovery))
colormap <- setNames(brewer.pal(length(methods)+1, "Set1"), c(methods, "NA"))
findings %>%
  mutate(method_of_discovery = factor(ifelse(is.na(method_of_discovery), "NA", method_of_discovery), levels=c(methods, "NA"))) %>%
  ggplot(aes(solve_status, fill=method_of_discovery)) +
  geom_bar() + scale_fill_manual(values=colormap, breaks=names(colormap))
```


# How many participants are pediatric?

```{r}
participant %>%
  mutate(pediatric = age_at_enrollment < 18) %>%
  count(pediatric)
```

Of these, how many are affected?

```{r}
participant %>%
  filter(affected_status %in% c("Affected", "Possibly affected")) %>%
  mutate(pediatric = age_at_enrollment < 18) %>%
  count(pediatric)
```

# How many short read genomes in total, and what percent are pediatric?

```{r}
sr <- participant %>%
  inner_join(experiment) %>%
  filter(table_name == "experiment_dna_short_read")
nrow(sr)
```

```{r}
sr %>% 
  mutate(pediatric = age_at_enrollment < 18) %>%
  count(pediatric)
```


# HPO frequency

```{r}
phenotype %>%
  filter(ontology == "HPO") %>%
  distinct(term_id) %>%
  nrow()
```

Top HPO terms

```{r}
phenotype %>%
  filter(ontology == "HPO") %>%
  count(term_id) %>%
  arrange(desc(n)) %>%
  filter(n > 100)
```

Do any probands not have HPO terms?

```{r}
hpo <- phenotype %>%
  filter(ontology == "HPO")
participants %>%
  filter(proband_relationship == "Self") %>%
  distinct(participant_id) %>%
  mutate(has_HPO = participant_id %in% hpo$participant_id) %>%
  count(has_HPO)
```

(Note same counts when considering all ontologies)
