---
title: Check files in WORKSPACES
author: Adrienne Stilp
date: July 10, 2023
editor_options: 
  chunk_output_type: console
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r install_packages}
install.packages(c(
  #"tidyverse",
  "kableExtra"
))
```

```{r libraries}
library(tools)
library(AnVIL)
library(tidyverse)
library(knitr)
```


```{r constants}
UPLOAD_CYCLE <- "U05"

# We may be able to automatically get this list somehow.
WORKSPACES_TO_KEEP <- c(
  ### U02
  "anvil-datastorage/AnVIL_GREGoR_GSS_U02_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U02_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U02_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U02_GRU",
  "anvil-datastorage/AnVIL_GREGoR_BCM_U02_GRU",
  "anvil-datastorage/AnVIL_GREGoR_CNH_I_U02_GRU",
  ### U03
  "anvil-datastorage/AnVIL_GREGoR_GSS_U03_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U03_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U03_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U03_GRU",
  "anvil-datastorage/AnVIL_GREGoR_BCM_U03_GRU",
  "anvil-datastorage/AnVIL_GREGoR_CNH_I_U03_GRU",
  ### U04
  "anvil-datastorage/AnVIL_GREGoR_GSS_U04_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U04_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U04_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U04_GRU",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U04_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BCM_U04_GRU",
  "anvil-datastorage/AnVIL_GREGoR_CNH_I_U04_GRU",
  ### U04
  "anvil-datastorage/AnVIL_GREGoR_GSS_U05_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U05_GRU",
  "anvil-datastorage/AnVIL_GREGoR_UW_CRDR_U05_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U05_GRU",
  "anvil-datastorage/AnVIL_GREGoR_BROAD_U05_HMB",
  "anvil-datastorage/AnVIL_GREGoR_BCM_U05_GRU",
  "anvil-datastorage/AnVIL_GREGoR_CNH_I_U05_GRU"
)
```

# Look up bucket for each workspace.

```{r}
WORKSPACES <- avworkspaces() %>%
  mutate(workspace=file.path(namespace, name)) %>%
  filter(workspace %in% WORKSPACES_TO_KEEP) %>%
  # Extract upload_cycle
  mutate(
    upload_cycle = str_extract(workspace, "_(U\\d+)", group=1)
  ) %>%
  # Add bucket path using avbucket
  rowwise() %>%
  mutate(
    path = avbucket(namespace=namespace, name=name)
  ) %>%
  select(workspace, namespace, name, upload_cycle, path)
WORKSPACES %>% select(workspace, upload_cycle, path) %>% kable()
```

# Pull out data tables from each workspace

```{r get_data_tables}
table_list <- list()
for (i in seq_along(WORKSPACES$workspace)) {
  workspace <- WORKSPACES$workspace[i]
  namespace <- WORKSPACES$namespace[i]
  name <- WORKSPACES$name[i]
  # Create a list for storing the tables.
  table_list[[workspace]] <- list()
  tables <- avtables(namespace=namespace, name=name)
  tmp_list <- list()
  for (table in tables$table) {
    tmp <- avtable(table, namespace=namespace, name=name) %>%
      # Only keep character type columns.
      select(where(is.character)) %>%
      # Add row numbers to have a key for the column.
      mutate(rownum=row_number()) %>%
      # Convert to long format.
      pivot_longer(!rownum, names_to="column", values_to="file") %>%
      # Only keep records where the value is gs.
      filter(str_detect(file, "^gs://")) %>%
      # no need to keep multiple rows with the same file
      select(-rownum) %>%
      distinct()
    tmp_list[[table]] <- tmp
    tmp <- bind_rows(tmp_list, .id="table")
    table_list[[i]] <- tmp
  }
}
table_data <- bind_rows(table_list, .id="referencing_workspace")
table_data %>% group_by(referencing_workspace) %>% count() %>% kable()
```

# Check files in all buckets

```{r get_files}
file_list <- list()
for (i in seq_along(WORKSPACES$workspace)) {
  bucket <- WORKSPACES$path[i]
  workspace <- WORKSPACES$workspace[i]
  
  top_level <- gsutil_ls(bucket, recursive=FALSE)
  subdirs <- top_level[str_detect(top_level, "/$")]
  # Don't check the notebooks and submissions directories.
  subdirs <- subdirs[!str_detect(subdirs, "notebooks|submissions")]
  # Using the ** makes the gsutil_ls command a lot faster.
  sub_levels <- lapply(subdirs, function(f) gsutil_ls(paste0(f, "**"), recursive=TRUE))
  tmp <- tibble(file=c(top_level, unlist(sub_levels)))
  
  file_list[[workspace]] <- tmp
}

file_data <- bind_rows(file_list, .id="workspace")
file_data %>% group_by(workspace) %>% count() %>% kable()
```

## Check if each file is referenced in a data table in any workspace

```{r}
tmp <- file_data %>%
  # Which tables are the files referenced in?
  left_join(table_data, by = "file") %>%
  # Which upload cycle are those WORKSPACES?
  left_join(WORKSPACES %>% select(workspace, upload_cycle), by=c("referencing_workspace"="workspace")) %>%
  select(workspace, file, referencing_workspace, upload_cycle) %>%
  arrange(workspace, file, upload_cycle, referencing_workspace)
```

```{r}
x <- tmp %>% 
  group_by(workspace, file) %>%
  summarise(
    n_workspaces=length(unique(na.omit(referencing_workspace))),
    n_upload_cycles=length(unique(na.omit(upload_cycle))),
    upload_cycles=paste(sort(unique(upload_cycle)), collapse=", "),
  ) %>%
  ungroup() %>%
  arrange(upload_cycles) %>%
  mutate(filetype = file_ext(file)) %>%
  filter(filetype != "") %>% # remove directories
  mutate(center=str_match(workspace, "anvil-datastorage/AnVIL_GREGoR_([A-Z_]+?)_U")[,2])

x0 <- x %>%
  filter(n_workspaces == 0) %>%
  select(-n_workspaces, -n_upload_cycles, -upload_cycles)
```

## Summary of files not referenced in any workspace by file type

```{r}
x0 %>%
  group_by(filetype, workspace) %>%
  count() %>%
  kable()
```

## List of files not referenced in any workspace by type {.tabset}

```{r, results="asis"}
for (category in unique(x$filetype)) {
  cat(sprintf("\n\n### %s\n\n", category))
    
  x0 %>%
    filter(filetype == !! category) %>%
    select(filetype, center, workspace, file, everything()) %>%
    arrange(filetype, center, workspace, file) %>%
    kable() %>%
    print()
}
```

## List of files not referenced in any workspace by RC {.tabset}

```{r, results="asis"}
for (category in unique(x$center)) {
  cat(sprintf("\n\n### %s\n\n", category))
    
  x0 %>%
    filter(center == !! category) %>%
    select(filetype, center, workspace, file, everything()) %>%
    arrange(filetype, center, workspace, file) %>%
    kable() %>%
    print()
}
```

## Save the list of files not referenced anywhere

```{r}
uri <- paste0(avbucket(), "/", UPLOAD_CYCLE, "_QC")
outfile <- paste(UPLOAD_CYCLE, "files_not_referenced_in_data_tables.tsv", sep="_")
write_tsv(x0, outfile)
# Copy to bucket
gsutil_cp(outfile, file.path(uri, outfile))
```

## Save a list of bam/cram/index files per RC

```{r}
for (category in unique(x0$center)) {
  outfile <- paste(category, UPLOAD_CYCLE, "files_not_referenced_in_data_tables.tsv", sep="_")
  x0_out <- x0 %>%
    filter(center == !! category) %>%
    filter(filetype %in% c("bam", "bai", "cram", "crai")) %>%
    arrange(workspace, file, filetype)
  if (nrow(x0_out) > 0) {
    write_tsv(x0_out, outfile)
    gsutil_cp(outfile, file.path(uri, outfile))
    # copy to workspaces so RCs can access
    rc_uri <- avbucket(namespace="anvil-datastorage", 
                       name=paste("AnVIL_GREGoR", category, UPLOAD_CYCLE, "GRU", sep="_"))
    gsutil_cp(outfile, file.path(uri, outfile))
    gsutil_cp(outfile, file.path(rc_uri, "post_upload_qc", outfile))
  }
}
```



