---
title: GREGoR Data Tracking - U1
author: DCC
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(AnVIL)
library(dplyr)
options("GCLOUD_SDK_PATH"="~/Applications/google-cloud-sdk") # don't run in AnVIL workspace
centers <- c("BCM", "CNH_I", "GSS", "BROAD", "UW_CRDR")
release <- "U1"
workspaces <- tibble(center=rep(centers, times=c(rep(1,4),2)),
                     consent=c(rep("GRU", 5), "HMB")) %>%
    mutate(workspace=paste("AnVIL_GREGoR", center, release, consent, sep="_"))
namespace <- "anvil-datastorage"
```

```{r}
participant <- lapply(workspaces$workspace, function(x) {
    tryCatch(
        avtable("participant", namespace=namespace, name=x) %>%
            select(participant_id, family_id) %>%
            mutate(workspace=x),
        warning=function(w) NULL, error=function(e) NULL)
}) %>% bind_rows() %>%
    left_join(workspaces)
```

```{r}
phenotype <- lapply(workspaces$workspace, function(x) {
    tryCatch(
        avtable("phenotype", namespace=namespace, name=x) %>%
            select(participant_id, phenotype_id) %>%
            distinct(participant_id, .keep_all=TRUE) %>%
            mutate(workspace=x),
        warning=function(w) NULL, error=function(e) NULL)
}) %>% bind_rows() %>%
    left_join(workspaces)
```

```{r}
analyte <- lapply(workspaces$workspace, function(x) {
    tryCatch(
        avtable("analyte", namespace=namespace, name=x) %>%
            select(analyte_id, participant_id, analyte_type) %>%
            mutate(workspace=x),
        warning=function(w) empty, error=function(e) empty)
}) %>% bind_rows() %>%
    left_join(workspaces)
```


```{r}
experiment_dna_short_read <- lapply(workspaces$workspace, function(x) {
    tryCatch(
        avtable("experiment_dna_short_read", namespace=namespace, name=x) %>%
            select(experiment_dna_short_read_id, analyte_id, experiment_type) %>%
            mutate(workspace=x),
        warning=function(w) NULL, error=function(e) NULL)
}) %>% bind_rows() %>%
    left_join(workspaces)
```

```{r}
aligned_dna_short_read <- lapply(workspaces$workspace, function(x) {
    tryCatch(
        avtable("aligned_dna_short_read", namespace=namespace, name=x) %>%
            select(aligned_dna_short_read_id, experiment_dna_short_read_id, aligned_dna_short_read_file) %>%
            filter(aligned_dna_short_read_file != "NA") %>%
            mutate(workspace=x),
        warning=function(w) NULL, error=function(e) NULL)
}) %>% bind_rows() %>%
    left_join(workspaces)
```


## Summary of Study Participants in AnVIL

```{r}
fam <- participant %>%
    group_by(center, consent) %>%
    summarise(participants=length(unique(participant_id)),
              families=length(unique(family_id)))

sum_fam <- fam %>%
    ungroup() %>%
    summarise(across(c(participants, families), sum)) %>%
    mutate(center="All", consent="All")

bind_rows(fam, sum_fam) %>%
    knitr::kable()
```


## Summary of sequencing experiments with files recorded in data table

```{r}
seq <- aligned_dna_short_read %>%
    distinct(center, consent, experiment_dna_short_read_id, .keep_all=TRUE) %>%
    count(center) %>%
    rename(`sequencing files`=n)

sum_seq <- seq %>%
    summarise(across(-center, sum)) %>%
    mutate(center="All")

bind_rows(seq, sum_seq) %>%
    knitr::kable()
```

## Summary of Data Completeness in AnVIL


```{r}
all_part <- participant %>%
    full_join(phenotype) %>%
    full_join(analyte) %>%
    full_join(experiment_dna_short_read) %>%
    full_join(aligned_dna_short_read) %>%
    distinct(center, consent, workspace, participant_id, .keep_all=TRUE)

ctr_part <- all_part %>%
    group_by(center) %>%
    summarise(has_pheno=sum(!is.na(phenotype_id)),
              has_seq=sum(!is.na(aligned_dna_short_read_id)),
              has_seq_pheno=sum(!is.na(aligned_dna_short_read_id) & !is.na(phenotype_id)),
              only_seq=sum(!is.na(aligned_dna_short_read_id) & is.na(phenotype_id)),
              only_pheno=sum(is.na(aligned_dna_short_read_id) & !is.na(phenotype_id)))

sum_part <- ctr_part %>%
    summarise(across(-center, sum)) %>%
    mutate(center="All")

bind_rows(ctr_part, sum_part) %>%
    knitr::kable()
```

- has_pheno: participants with phenotype data in AnVIL
- has_seq: participants with sequencing data files in AnVIL
- has_seq_pheno: participants with both sequencing and phenotype data
- only_seq: participants with sequencing data but without phenotype data
- only_pheno: participants with phenotype data but without sequencing data
